Debiased Calibration for Generalized Two-Phase Sampling

### 2025-01-23 final/paper/paper.tex
Updated paper with Dr. Kim's suggestions.

Updates:
\begin{itemize}
  \item Modified the end of paragraph two of the introduction and the beginning
    of paragraph three of the introduction to mirror more of the content of the
    rest of the paper.
  \item Fixed spacing issues with the beginning of paragraphs.
  \item Added Table~\ref{tab:msexample}
\end{itemize}
\newpage

@@

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
second-phase sample. From the first-phase sample, we can compute
$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
We therefore have three estimators for
two parameters. The two-phase regression estimator can be  used to combine this
information efficiently and it can be expressed as  
\begin{equation} \label{eq:1}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\end{equation} 
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and

$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
 \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
$$

Regarding the choice of $q_i$, one can consider a linear regression model 
\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\label{regression}
\end{equation}
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
function. In this case,  $q_i = c_i$ is a reasonable choice. 
One way to understand the two-phase regression estimator is to use the following
decomposition:

$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{\bf x_i^\top \hat{\bm \beta_q}}{\pi_{1i}}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor
under the working superpopulation model. The bias correction term 
makes the two-phase regression estimator asymptotically design unbiased
regardless of the underlying the superpopulation model. 
Note that the two-phase regression estimator in Equation~\ref{eq:1} can be
expressed as  $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} \pi_{1i}^{-1} \hat w_{2i} y_i$ where

$$
\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
\bf x_i / \pi_{1i}.
$$

This means that $\hat Y_{\rm reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ being continuous.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others \citep{schennach2007point}.

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. We would rather
have bias calibration separate from the minimizing the variance so that
we can control each in isolation. 
\cite{kwon2024debiased} shows that for a generalized entropy
function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

@@

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
strictly convex and differentiable. We consider maximizing 
\begin{equation}\label{eq:primalloss}
  H(w_2) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
\end{equation}
subject to the following constraints:
\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
\end{equation}
and 
\begin{equation}\label{eq:calconst2}
  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}, 
\end{equation}
where $g(w) = \partial G / \partial w$.

The first constraint is the usual calibration constraint motivated from the
regression model in (\ref{regression}) and the second
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}.
By writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
\end{equation}
where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
estimator of $Y_N$ is 
\begin{equation}\label{eq:tpest}
\hat Y_{\rm DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i.
\end{equation}

@@

We can understand the estimated weights, $\hat w_{2i}$ from
Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
minimize the Lagrangian function
\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  + \bm \lambda^\top \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
\end{equation}
where $\bm \lambda$ is a vector of Lagrange multipliers.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, we can obtain 
$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
where $g(w) = \partial G / \partial w$ and 
$\hat{\bm \lambda}$ is the solution to
\begin{equation}\label{eq:lamdc1}
  \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
  \sum_{i \in A_2} w_{1i} w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
\end{equation}
The following theorem shows that the construction of the weights in
Equation~\ref{eq:primal} leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
  Under the regularity conditions stated in the appendix
  $$
  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
  $$
  where
  $$
  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\bm \phi^*,
  $$
  and
  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
  $$

%   \begin{align*}
%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
%  \end{align*} 
%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
%  and  
%
%  $$
%  \begin{pmatrix}
%  \bm \phi_1^* \\
%  \phi_2^* 
%  \end{pmatrix}
%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
%  \begin{pmatrix}
%  \bx_i \bx_i^\top &   \bx_i g_i   \\
%  g_i  \bx_i^\top   & g_i^2     
%  \end{pmatrix} \right]^{-1}
%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
%  \begin{pmatrix}
%    \bx_i \\ g_i 
%  \end{pmatrix}y_i 
%  $$

@@

See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
theorem implies
that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
the same asymptotic expectation and variance, 
which means that we can estimate
the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
method is equivalent to a regression estimator. By constraints 
(\ref{eq:calconst1}) and (\ref{eq:calconst2}), 
$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent, and the
variance is asymptotically equivalent to the two-phase regression estimator. 
We can estimate $\bm \phi^*$ with
\begin{equation}
\hat{\bm \phi} = 
\left[\sum_{i \in A_2} 
w_{1i}\frac{\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
\label{eq:8}
\end{equation}
Therefore, the proposed calibration estimator in (\ref{eq:primal}) is
asymptotically equivalent to the two-phase regression estimator given by 
  $$
  \hat Y_{\rm tp, reg}  = 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\hat{\bm \phi},
  $$
where $\hat{\bm \phi}$ is defined in (\ref{eq:8}).

For specific sample designs, we can construct the optimal entropy function
$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
in $\bm \phi^*$. Consider a class of generalized regression estimators,
$$
\mathbf{B} = 
\left\{\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \gamma + 
\sum_{i \in A_2} \frac{w_{1i}}{\pi_{2i|1}}(y_i - \bf{z}_i^\top \bm \gamma):
\bm \gamma \in \R^{p+1}\right\}
$$
Then the design optimal estimator is 
$$
\hat \theta_{\ell, \rm{opt}} = 
\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
\frac{w_{1i}}{\pi_{2i|1}}\sum_{i \in A_2} (y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
$$
where 
$$
\bm \phi_{\rm opt} = 
\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
$$
We assume that both $A_1$ and $A_2$ are Poisson samples. Then,
$$
\bm \phi_{\rm opt} = 
\left\{
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
\right\}^{-1}
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
$$
To find the optimal entropy function, we solve the differential equation
$$
\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
$$
Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
entropy function is 
$$
G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}w_{1i} - 1).
$$
To construct an asymptotically unbiased estimate of the variance we assume that
the sampling designs for $A_1$ and $A_2$ are measurable and we define the
following joint inclusion probabilities:
$$
\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
$$
The next theorem discusses how to compute an estimate for the variance of 
$\hat Y_{\rm DCE}$. For the proof see the appendix.

@@

  Define $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
  $$
  \hat V_{\rm DCE} 
  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
  \hat \eta_i \hat \eta_j
  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}
  $$
  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.
\end{theorem}

@@

Now we consider the sampling mechanism known as non-nested two-phase sampling 
\citep{hidiroglou2001double}. In the last section, we considered two-phase sampling
in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
consist of observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample,
$A_2$, to consist of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.

>>>>

@@

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
second-phase sample. We write $d_{1i} = \pi_{1i}^{-1}$ and $d_{2i|1} =
\pi_{2i|1}^{-1}$ to denote the design weights.
From the first-phase sample, we can compute
$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} d_{1i}\bx_i$ as an unbiased
estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
\sum_{i \in A_2} d_{1i} d_{2i \mid 1} \left( \bx_i, y_i\right)$ as
an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
We therefore have three estimators for
two parameters. The two-phase regression estimator can be  used to combine this
information efficiently and it can be expressed as  
\begin{equation} \label{eq:1}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} d_{1i} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} d_{1i} d_{2i|1} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\end{equation} 
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
d_{1i} \bf x_i \bf x_i^\top q_i \right)^{-1} 
 \sum_{i \in A_2} d_{1i}\bf{x}_i y_i q_i.
$$

Regarding the choice of $q_i$, one can consider a linear regression model 
\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\label{regression}
\end{equation}
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
function. In this case,  $q_i = c_i^{-1}$ is a reasonable choice. 
One way to understand the two-phase regression estimator in \eqref{eq:1}
is to use the following decomposition: 
$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} d_{1i}\bf x_i^\top \hat{\bm \beta_q}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
d_{1i} d_{2i|1}(y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor
under the working superpopulation model. The bias correction term 
makes the two-phase regression estimator asymptotically design unbiased
regardless if the underlying the superpopulation model is correct or not.
We can find an equivalent estimator in the form $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} d_{1i} \hat w_{2i} y_i$. This is useful because it shows that
the estimator is linear and it is interesting because the estimated weights
$\hat w_{2i}$ are the weights closest to the design weights that satisfy the
constraints from \eqref{eq:constr}.
\begin{equation}\label{eq:constr}
\sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i 
= \sum_{i \in A_1} d_{1i} \bf{x}_i.
\end{equation}
Closeness is in an $L^2$-sense. Overall, the weights $\hat w_{2i}$ are the
solution to the following optimization problem:
$$
\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - d_{2i|1})^2 q_i
\text{ such that } \sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i 
= \sum_{i \in A_1} d_{1i} \bf{x}_i.
$$

This means that $\hat Y_{\rm reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ being continuous with respect 
to $w_i$.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others \citep{schennach2007point}.

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. We would rather
have bias calibration separate from the minimizing the variance so that
we can control each in isolation. 
\cite{kwon2024debiased} shows that for a generalized entropy
function $G(w)$, including a term of $g(d_{2i|1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

@@

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
strictly convex and differentiable. We consider maximizing 
\begin{equation}\label{eq:primalloss}
  H(w_2) = - \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i
\end{equation}
subject to the following constraints:
\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i = 
  \sum_{i \in A_1} d_{1i}\bf{x}_i
\end{equation}
and 
\begin{equation}\label{eq:calconst2}
  \sum_{i \in A_2} d_{1i} w_{2i} g(d_{2i|1})q_i = 
  \sum_{i \in A_1} d_{1i} g(d_{2i|1})q_i, 
\end{equation}
where $g(w) = \frac{\partial G}{\partial w}$.

Also known as covariate balancing \citep{imai2014covariate},
constraint~\eqref{eq:calconst1} ensures that the weighted mean of the
covariates, $\bf{x}_i$, are the same.
This is also the usual calibration constraint motivated from
the regression model in \eqref{regression}.
Constraint~\eqref{eq:calconst2} is how we incorporate the design weights. Using
$g(w) = \frac{\partial G}{\partial w}$, ultimately ensures that $\hat w_{2i}$
converges to $d_{2i}$ in probability. This 
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}. We extend
their method within our two-phase framework to allow for an estimated quantity
on the right hand side of \eqref{eq:calconst1}.
To estimate the weights, we solve
\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i = \sum_{i \in A_1} d_{1i} \bf z_i 
\end{equation}
where  $\bf z_i = (\bf x_i , g(d_{2i|1})q_i)$. The resulting 
estimator of $\bar Y_N$ is 
\begin{equation}\label{eq:tpest}
  \hat Y_{\rm DCE} = N^{-1} \sum_{i \in A_2} d_{1i} \hat w_{2i} y_i.
\end{equation}

@@

We compute the estimated weights, $\hat w_{2i}$ from
\eqref{eq:primal} using the method of Lagrange multipliers. We 
minimize the Lagrangian function 
\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i 
  + \bm \lambda^\top \left( \sum_{i \in A_1} d_{1i} \bf z_i -
    \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i\right)
\end{equation}
where $\bm \lambda \in \R^{p+1}$ is a vector of Lagrange multipliers and
$\bf{z}_i = (\bf{x}_i^\top, g(d_{2i|1})q_i)^\top$.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, we can obtain 
$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
where $g(w) = \frac{\partial G}{\partial w}$ and 
$\hat{\bm \lambda}$ is the solution to
\begin{equation}\label{eq:lamdc1}
  \sum_{i \in A_1} d_{1i} \bf{z}_i  -
  \sum_{i \in A_2} d_{1i} w_{2i}(\hat{\bm \lambda}) \bf{z}_i = 0.
\end{equation}
Under regularity conditions presented in the appendix, this solution is unique.
The following theorem presents a linearization of
\eqref{eq:primal} which leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in \eqref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
  Under the regularity conditions stated in the appendix
  $$
  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
  $$
  where
  $$
  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_2} d_{1i} d_{2i|1} y_i
  + \left(\sum_{i \in A_1} d_{1i} \bf z_i - 
  \sum_{i \in A_2} d_{1i} d_{2i|1} \bf{z}_i \right)\bm \phi^*,
  $$
  and
  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top}{g'(d_{2i|1})q_i}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})q_i}.
  $$

%   \begin{align*}
%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} d_{1i} \bf x_i  -
%     \sum_{i \in A_2} d_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
%   &\quad+ \left(\sum_{i \in A_1} d_{1i} g_i  -
%  \sum_{i \in A_2} d_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
%  \end{align*} 
%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
%  and  
%
%  $$
%  \begin{pmatrix}
%  \bm \phi_1^* \\
%  \phi_2^* 
%  \end{pmatrix}
%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
%  \begin{pmatrix}
%  \bx_i \bx_i^\top &   \bx_i g_i   \\
%  g_i  \bx_i^\top   & g_i^2     
%  \end{pmatrix} \right]^{-1}
%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
%  \begin{pmatrix}
%    \bx_i \\ g_i 
%  \end{pmatrix}y_i 
%  $$

@@

See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
theorem implies
that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
the same asymptotic expectation and variance, 
which means that we can estimate
the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
method is equivalent to a two-phase regression estimator. By Theorem~\ref{thm:dc1}, 
$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent,
and the proposed calibration estimator in \eqref{eq:primal} is
asymptotically equivalent to the two-phase regression estimator given by 
  $$
  \hat Y_{\rm tp, reg}  = 
  \sum_{i \in A_2} d_{1i} d_{2i|1} y_i
  + \left(\sum_{i \in A_1} d_{1i} \bf z_i  - 
  \sum_{i \in A_2} d_{1i} d_{2i|1} \bf{z}_i \right)\hat{\bm \phi},
  $$
where $\hat{\bm \phi}$ is defined in \eqref{eq:8}
\begin{equation}
\hat{\bm \phi} = 
\left[\sum_{i \in A_2} 
d_{1i}\frac{\bf z_i \bf z_i^\top}{g'(d_{2i|1})q_i}\right]^{-1}
\sum_{i \in A_2} d_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})q_i}.
\label{eq:8}
\end{equation}

For specific sample designs, we can construct the optimal entropy function
$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
in $\bm \phi^*$. Consider a class of generalized regression estimators,
$$
\mathbf{B} = 
\left\{\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \gamma + 
\sum_{i \in A_2} d_{1i}d_{2i|1}(y_i - \bf{z}_i^\top \bm \gamma):
\bm \gamma \in \R^{p+1}\right\}
$$
The design optimal estimator is 
$$
\hat \theta_{\ell, \rm{opt}} = 
\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
\sum_{i \in A_2} d_{1i}d_{2i|1}(y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
$$
where 
$$
\bm \phi_{\rm opt} = 
\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
$$
We assume that both $A_1$ and $A_2$ are Poisson samples. Then,
$$
\bm \phi_{\rm opt} = 
\left\{
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
\right\}^{-1}
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
$$
To find the optimal entropy function, we solve the differential equation
$$
\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
$$
Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
entropy function is 
$$
G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}d_{1i} - 1).
$$
To construct an asymptotically unbiased estimate of the variance we assume that
the sampling designs for $A_1$ and $A_2$ are measurable and we define the
following joint inclusion probabilities:
$$
\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
$$
The next theorem discusses how to compute an estimate for the variance of 
$\hat Y_{\rm DCE}$. For the proof see the appendix. The idea behind the proof is
to use the two-phase sampling approach to estimate the variance. If the sampling
fraction $n_2 / n_1 \to 0$ as $N$ increases, then the reverse framework of
\cite{fay1992inferences} allows us to make a good approximation.

@@

  Define $\hat \eta_i = \bf z_i \hat{\bm \phi} + 
  \delta_i d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})$.
  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
  $$
  \hat V_{\rm DCE} 
  = \sum_{i \in A_1} \sum_{j \in A_1} d_{1ij} \Delta_{1ij}
  \hat \eta_i \hat \eta_j
  + \sum_{i \in A_2} \sum_{j \in A_2} d_{1i} d_{2ij|1}\Delta_{2ij|1}
  d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})
  d_{2j|1}(y_j - \bf z_j \hat{\bm \phi})
  $$
  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$,
  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
  $d_{1ij} = \pi_{1ij}^{-1}$, $d_{2ij|1} = \pi_{2ij|1}^{-1}$,
  $\pi_{1ij} = \Pr(i \in A_1, j \in A_1)$, and 
  $\pi_{1ij} = \Pr(i \in A_2, j \in A_2 | i \in A_1, j \in A_1)$.
\end{theorem}

@@

We now consider the sampling mechanism known as non-nested two-phase sampling 
\citep{hidiroglou2001double}. In the last section, we considered two-phase sampling
in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
consist of observations of $(\bf X_i)$ and the Phase 2 sample,
$A_2$, to consist of observations of $(\bf X_i, Y_i)$. We denote the sample size
of $A_1$ as $n_1$ and the sample size of $A_2$ as $n_2$.
